---
title: "RStudio on SageMaker User Counting"
format: gfm
editor: source
---

This demonstrates how to use the PAWs R package to query SageMaker user details and determine how many users have access to RStudio on SageMaker. While this doesn't necessarily directly detail how many users **have** accessed RStudio, it does provide detail on how many users **can** access RStudio. This is helpful when multiple SageMaker domains are setup and you want to know how many users have access to RStudio across all domains.

### Setup
Load the `paws` package.
```{r}
library(paws)
```

Create the `sagemaker` client. This is where you specify the profile / credentials to use. In this case, we're using the `example` profile defined in our [AWS config file](https://docs.aws.amazon.com/cli/latest/userguide/cli-configure-files.html).
```{r}
sagemaker <- paws::sagemaker(credentials = list(profile = "pm-account"))
```

### User Counting
Query the list of users using `$list_user_profiles()`:
```{r}
users <- sagemaker$list_user_profiles()
```

In order to retrieve user details, we need the `UserProfileName` and `DomainId` for each user. We'll iterate over the `users` list and use `$describe_usr_profile()` to get the details for each user:
```{r}
all_user_details <- lapply(users$UserProfiles, function(profile) {
  sagemaker$describe_user_profile(UserProfileName = profile$UserProfileName, DomainId = profile$DomainId)
})
```

Using `user_details`, we can use `UserSettings` to determine if the user has access to RStudio. First, we'll define a function to extract and format details about user RStudio access:
```{r}
extract_rstudio_status <- function(user_details) {
  domain_id = user_details$DomainId
  user_profile_name = user_details$UserProfileName
  user_profile_arn = user_details$UserProfileArn
  rstudio_status = user_details$UserSettings$RStudioServerProAppSettings$AccessStatus
  rstudio_usergroup = user_details$UserSettings$RStudioServerProAppSettings$UserGroup

  data.frame(
    domain_id = domain_id,
    user_profile_name = user_profile_name,
    user_profile_arn = user_profile_arn,
    rstudio_status = ifelse(length(rstudio_status) == 0, NA, rstudio_status),
    rstudio_usergroup = ifelse(length(rstudio_usergroup) == 0, NA, rstudio_usergroup)
  )
}
```

We can then use this function to extract RStudio access details for each user and combine them into a single data frame:
```{r}
reduce_rbind <- \(x) Reduce(rbind, x)

sm_user_rstudio_status <- lapply(all_user_details, extract_rstudio_status) |> 
  reduce_rbind()

tibble::as_tibble(sm_user_rstudio_status)
```

This final data frame will contain details about each user's RStudio access status. You can then use this data frame to determine how many users across each domain have access to RStudio on SageMaker. Again, this doesn't necessarily detail how many users **have** accessed RStudio, but how many users **can** access RStudio.